---
title: "Weekly Notes"
subtitle: "Course DS 350"
author: "Bethany Ball"
format:
  html:
    self-contained: true
    page-layout: full
    title-block-banner: true
    toc: true
    toc-depth: 3
    toc-location: body
    number-sections: false
    html-math-method: katex
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-copy: hover
    code-tools:
        source: false
        toggle: true
        caption: See code
execute: 
  warning: false
    
editor_options: 
  chunk_output_type: console
---

### Weekly Notes

# section heading (write with on # first)
text text text (just writing plain text)
**bold** or __bold__ (2 starts or 2 underscores before and after the word(S))
_italics_ (italics underscore before and after)

below is how to make sure that that the code is in R.
...{r}

...

short-key for a new code chunk is :
PC: ctrl_alt_i
Mac command + option +i

### GGPlot
#install.packages("tidyverse")
library(tidyverse)

mtcars

head(mtcars)
mtcars %>% head()
# %>% (short key: command + Shift + m)

tail(mtcars)
glimpse(mtcars)

view(mtcars)






```{r}
library(tidyverse)
```


```{r}
mtcars %>% view()
```


```{r}
ggplot(data = mtcars, mapping = aes(x = factor(cyl), y = mpg, color = cyl)) + 
  geom_point() +
  geom_violin() +
  geom_line()
```


```{r}
install.packages('nycflights13')
```

```{r}
library(nycflights13)

#flights %>% view()

ggplot(data = flights, mapping = aes(x=arr_time, y = sched_arr_time)) +
  geom_point()

ggplot(data = flights, aes(x = dep_time, y = dep_delay, color =
  (dep_delay<0)))+
  geom_point()

 

```


















