```{r}
library(tidyverse)

```

```{r}
#install.packages("nycflights13")
data <- nycflights13::flights

```

### Week 2: Task 1

**Task: Visualizing Large Distributions**

```{r}
# Univariate distribution for dep_delay
ggplot(data, aes(x = dep_delay)) +
  geom_density(fill = "blue", alpha = 0.5) +
  labs(title = "Density Plot of Departure Delays", x = "Departure Delay (minutes)", y = "Density")

# Univariate distribution for arr_delay
ggplot(data, aes(x = arr_delay)) +
  geom_density(fill = "red", alpha = 0.5) +
  labs(title = "Density Plot of Arrival Delays", x = "Arrival Delay (minutes)", y = "Density")


```

The two charts above respectively display the distributions of *arrival*
delays and *departure* delays. At first glance, they may seem almost
identical, being different only in the color. However, they represent
distinct data points. While both distributions share similarities in
their overall shape, they are different.

The chart below shows both variables in a single visualization,
providing a more complete picture of the relationship between departure
and arrival delays. It shows a strong linear correlation, showing that
flights that depart late usually arrive late. However, the data also
show that the delays are not always the same this shows that departure
and arrival delays are not simply the same data viewed from different
lens.

These visualizations show that, as a general rule, late departures lead
to late arrivals. However, they also show that a flight can still arrive
late even when it departs on time, typically by no more than an hour, or
so. This suggests that other factors such as weather or air traffic can
effect a flight’s arrival time, even when departure is on time Overall,
the findings align with expectations: late departures often cause late
arrivals, but other factors can contribute to delays, regardless of
departure timing.

```{r}
#install.packages("hexbin")
#library(hexbin)

#ggplot(data, aes(x = dep_delay, y = arr_delay)) +
#  geom_hex(bins = 50) +
#  scale_fill_viridis_c() +
#  labs(title = "Hexbin Plot of Departure and Arrival Delays", 
#      x = "Departure Delay (minutes)", y = "Arrival Delay (minutes)") +
#  theme_minimal()

#```

#```{r}
# Load necessary libraries
library(ggplot2)

# Create a point plot
ggplot(data, aes(x = dep_delay, y = arr_delay)) +
  geom_point(alpha = 0.5, color = "blue", size = 1) +
  labs(title = "Point Plot of Departure and Arrival Delays", 
       x = "Departure Delay (minutes)", y = "Arrival Delay (minutes)") +
  theme_minimal()

```
