---
title: "Task: Country Heights"
subtitle: "Course DS 350"
author: "Bethany Ball"
keep-md: true
fig-width: 12  
fig-height: 10
format:
  html:
    self-contained: true
    page-layout: full
    title-block-banner: true
    toc: true
    toc-depth: 3
    toc-location: body
    number-sections: false
    html-math-method: katex
    code-fold: true
    code-summary: "Show the code"
    code-overflow: wrap
    code-copy: hover
    code-tools:
        source: false
        toggle: true
        caption: See code
execute: 
  warning: false
    
editor_options: 
  chunk_output_type: console
---


```{r}
library(haven)
library(foreign)
library(readr)
library(dplyr)
library(ggplot2)
library(foreign)

#library(pacman)

#pacman ::(haven, foreign, readr, dplyr, ggplot2, foreign, tidyverse, rio, pander)

```

```{r}

german_conscr = read_dta("germanconscr.dta")
# bdec - birth year
# "height" - cm
# all M
german_conscr

german_prison = read_dta("germanprison.dta")
#"bdec" - birth year
#"height" - cm
# all M
german_prison

Heights_south_east <- read.dbf("Heights_south-east/B6090.dbf")#Birth Year: "ALTER" 
#Height: "CMETER"
#sex: "F" 5 = m, 6 = f
head(Heights_south_east)

BLS = read_csv("https://raw.githubusercontent.com/hadley/r4ds/main/data/heights.csv")
#"height" in
#sex (male, female)
#birthday all: 1950
BLS

UWNSD = read_sav("main05022005.sav")
#RT216I height of respondent, inches
#doby DATE OF BIRTH, YEAR
#RE35 male = 1, female = 2, other(drop/na)
UWNSD


# colnames(Heights_south_east)

```




```{r}
# Wrangle datasets
german_conscr <- german_conscr %>%
  mutate(birth_year = bdec, height.cm = height, height.in = height / 2.54, study = "German Conscripts") %>%
  select(birth_year, height.in, height.cm, study)

german_prison <- german_prison %>%
  mutate(birth_year = bdec, height.cm = height, height.in = height / 2.54, study = "German Prison") %>%
  select(birth_year, height.in, height.cm, study)
```


```{r}
library(dplyr)

# Update Heights_south_east dataset
Heights_south_east <- Heights_south_east %>%
  mutate(
    birth_year = 1800 + ALTER,  
    height.cm = CMETER, 
    height.in = CMETER / 2.54, 
    study = "Heights South-East"
  ) %>%
  select(birth_year, height.in, height.cm, study)

head(Heights_south_east)

```


```{r}
BLS <- BLS %>%
  filter(sex == "male") %>%
  mutate(birth_year = 1950, height.in = height, height.cm = height * 2.54, study = "BLS") %>%
  select(birth_year, height.in, height.cm, study)
```

```{r}

# RT216F (feet)
# RT216I (inches)
UWNSD_FI <- UWNSD %>% 
  filter(
    RT216F != "-1" & RT216F != "-2" & RT216F != "." &
    RT216I != "-1" & RT216I != "-2" & RT216I != "25" & 
    RT216I != "35" & RT216I != "45" & RT216I != "50" & RT216I != "85"
  ) %>%
  mutate(height_in = as.numeric(RT216F) * 12 + as.numeric(RT216I))



```


```{r}
library(dplyr)

UWNSD <- UWNSD_FI %>%
  mutate(
    birth_year = DOBY,
    height.in = height_in,
    height.cm = height_in * 2.54,  
    study = "UWNSD"
  ) %>%
  select(birth_year, height.in, height.cm, study)

#print(UWNSD)

# include the full year
UWNSD <- UWNSD %>%
  mutate(
    birth_year = ifelse(birth_year <= 5, 2000 + birth_year, 1900 + birth_year)
  )

print(UWNSD)

```


```{r}
# Combine datasets
combined_data <- bind_rows(german_conscr, german_prison, Heights_south_east, BLS, UWNSD)

library(ggplot2)
library(plotly)

plot_1 <- ggplot(combined_data, aes(x = birth_year, y = height.cm, color = study)) +
  geom_point() +
  labs(title = "Height Distribution Across Studies",
       x = "Birth Year", y = "Height (cm)") +
  theme_minimal()

ggplotly(plot_1)



```


```{r}
german_conscr
german_prison
head(Heights_south_east)
BLS
UWNSD

combined_data
```


Paragraph on Data Wrangling

The data wrangling process involved importing five datasets in various formats, including Stata, DBF, CSV, and SPSS. Each dataset required careful cleaning, standardizing column names, and converting heights between inches and centimeters. For datasets that did not have a clear birth year, assumptions were made (e.g., for the BLS data, a birth year of 1950 was assumed). One challenge was identifying the birth year in the German soldier data, where column names were in German. After using Google Translate, the correct column (geburtsjahr) was identified. Overall, I had to ensure consistency in column names (birth_year, height.in, height.cm, and study) before combining the datasets.


Analysis and Comparison

The data suggests differences in average heights across birth years and geographic regions. While the datasets show an increase in height over time, especially in 20th-century data, there are variations. For example, the German conscripts from the 19th century appear shorter than those in the mid-20th century BLS data. This aligns with historical trends showing a gradual increase in human height due to better nutrition, health, and living conditions.

In comparison to the previous task, this analysis confirms the trend that humans have become taller over the centuries, but the rate of growth and regional differences highlight the complexity of height distribution. Some discrepancies could arise due to differences in data collection methods, sample populations, or societal conditions during the studied periods.
Conclusion

Overall, the data supports the notion that humans have been getting taller over time, though the extent of height increase varies based on geography and historical context. This increase likely results from improved nutrition and health care over time. However, it’s important to consider the limitations of historical datasets and the potential biases that may affect this conclusion.











